---
title: "CS 422 Section 04 Homework 1"
author: "Thomas Ehling A20432671"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

## 2 Problems

```{r}
library(ggplot2)
library(dplyr)
library(corrplot)
library(psych)
library(textreuse)
library(stringr)
```


## 2.1 Problem 1

###2.1.A Visualizing the dataset

```{r}
setwd('/home/french_magician/Documents/Chicago/DM/Assigment_4/')
files <- list.files("hw4.movies", full.names=T)
corpus <- TextReuseCorpus(files, tokenizer = tokenize_ngrams, n = 5,keep_tokens = TRUE)
```
```{r}
#Number of documents
cat(paste("the number of columns is ",length(names(corpus))))
```

```{r}
# number of shingles
cat(paste("the number of rows is ",length(tokens(corpus))))
```

###2.1.B Visualizing the dataset

```{r}
d <- corpus[["user20"]]
names(d)
```

###2.1.B-i

```{r}
# with str count we need to select keywords to look for, 
# I selected the 3 first digits of the year, from 1900 to 2019
cat(paste("the number of rated movies is ",str_count(content(d), '190|191|190|193|194|195|196|197|198|199|200|201')))
```

###2.1.B-ii

```{r}
cat(paste("the five fist shingles are : \n"))
head(tokens(d),5)
```

###2.1.C

```{r}
comparisons <- pairwise_compare(corpus, jaccard_similarity, progress = TRUE)
```

```{r}
candidates <- pairwise_candidates(comparisons)
names(candidates)
```
###2.1.C-i

```{r}
cat(paste("there are ",length(which(candidates$score>0.6))," candidates with a score > 0.6"))
```

###2.1.C-ii

```{r}
cat(paste("there are ",length(which(candidates$score>0.5))," candidates with a score > 0.5"))
```

###2.1.C-iii

```{r}
cat(paste("there are ",length(which(candidates$score>0.4))," candidates with a score > 0.4"))
```

###2.1.C-iv

```{r}
cat(paste("the candidates with a score > 0.4 are : \n"))
candidates[candidates$score > 0.4, ]
```

###2.1.D-i

```{r}
lsh_probability(h=18,  b=9, s=0.6)
```
For a jaccar similarity of 0.6 and a probability of 0.98, the lowest number of minhashes is 18 and bands is 9.

###2.1.D-ii

```{r}
minhash <- minhash_generator(n=18, seed=100)
corpus <- TextReuseCorpus(files, tokenizer = tokenize_ngrams, n = 5,
                          minhash_func = minhash, keep_tokens = TRUE)
```

```{r}
doc <- corpus[["user20"]]
tokens(doc)[1:5]
```

###2.1.E


```{r}
buckets <- lsh(corpus, bands = 9)
candidates <- lsh_candidates(buckets)
res <- lsh_compare(candidates, corpus, jaccard_similarity)
```

###2.1.E-i

I don't expect to find any, because we didn't find any before

###2.1.E-ii

```{r}
names(res)
```


```{r}
cat(paste("the candidates with a score > 0.4 are : \n"))
res[res$score > 0.4, ]
```

There ae 3 observations.

###2.1.E-iii

There are exactly the same one as above

###2.1.E-iv

In c we mase 224,115 comparisons, here we made 1,026 comparisons.

So we saved (224,115 - 1026)/224,115*100 = 99.5% of the work.

###2.2 Topic: Content-based recommendation system

My AID is A23462671, so my userId is 234671 mod 671 = 50.

```{r}
df_movies <- read.csv("ml-latest-small/movies.csv", sep=",", header=T)
head(df_movies,6)
```

```{r}
df_user_50 <- read.delim("hw4.movies/user50.txt", header = FALSE, sep = "\t")
tail(df_user_50)
nb_movies = nrow(df_user_50)
```

```{r}

gendra_list <- c("Action", "Adventure", "Animation", "Children", "Comedy", "Crime", "Documentary", "Drama", "Fantasy",
"Film-Noir", "Horror", "IMAX", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller", "War", "Western", "(no
genres listed)")

nb_gendras = length(gendra_list)

df_movie_profile = data.frame(matrix(vector(), 0, 20,
                dimnames=list(c(), gendra_list)), stringsAsFactors=F)

df_gendras = df_user_50

for (ind_gendra in 1:nb_gendras){
  
  list_gendra <- character(nb_movies)
  gendra <- gendra_list[ind_gendra]
  
  for (ind_movie in 1:nb_movies){
  
    title <- paste(df_user_50$V1[ind_movie])
    movie <- df_movies[which(df_movies$title == title),]
    gendras <- paste(movie$genres)
    gendras <- strsplit(gendras,"|", fixed=TRUE)
    my_gendras <- gendras[[1]]

    if(gendra %in% my_gendras){
      list_gendra[ind_movie] <- 1
    }else{
      list_gendra[ind_movie] <- 0
    }
  }
  df_gendras[,gendra] <- list_gendra
}

df_gendras
```

```{r}

profile_user_50 <- integer(nb_gendras)

for(j in 1:nb_gendras+1){

  sum_val <- 0
  
  for(i in 1:nb_movies){
    sum_val <- sum_val + as.numeric(df_gendras[i,j])
  }
  
  profile_user_50[j-1] <- round(sum_val/nb_movies, digits = 2)

}


profile_user_50

```


```{r}
set.seed(1122)
random_movies <- sample(df_movies$movieId, 10 )
random_movies
```

```{r}
get_movie_profile <- function(movie_id) {
  
  list_actions =  integer(nb_gendras)
  
  movie <- df_movies[which(df_movies$movieId == movie_id),]
  
  gendras <- paste(movie$genres)
  
  gendras <- strsplit(gendras,"|", fixed=TRUE)
  
  my_gendras <- gendras[[1]]
  
  for(i in 1:nb_gendras){
    if(gendra_list[i] %in% my_gendras){
        list_actions[i] <- 1
    }else{
        list_actions[i] <- 0
    }
  }
  
  return(list_actions)
  
}

length(get_movie_profile(145))

```

```{r}
my_cosine <- function(x, y) {
  sum(x*y)/(norm(x, type ="2") * norm(y, type = "2"))
}
```

```{r}

# Calculate the sin values
sin_values <- integer(nb_movies)
for(i in 1:length(random_movies)){
  sin_values[i] <- my_cosine(profile_user_50, get_movie_profile(random_movies[i]))
}


# create a new sorted list to spot the best ones
sorted_sin <- sort(sin_values, decreasing = TRUE)
#head(sorted_sin,5)

# remove duplicate vaules among the sorted file
sorted_sin <- sorted_sin[!duplicated(sorted_sin)]
#head(sorted_sin,5)

# create a list with the sorted indexes of movies
indexes <- c()
for(i in 1:5){
  indexes <-c(indexes, which(sin_values==sorted_sin[i]))
}

cat(paste("User Id 50 chose the following 10 movies : ",paste(random_movies,collapse=" "),"\n\n"))
cat(paste("Of these, the following 5 are recommended : \n\n"))
# get the 5 best movies from the indexes and print it :
cat(paste("MovieId  MovieName Similarity \n\n"))
for(i in 1:5){
  
  id_rd_movie <-random_movies[indexes[i]]
  
  id_movie <- which(df_movies$movieId == id_rd_movie)
  
  movie_name <- df_movies$title[id_movie]
  
  similarity <- sin_values[indexes[i]]
  
  cat(paste(id_rd_movie," ",movie_name," ",similarity,"\n\n"))
  
}


```

###2.3 Topic: Collaborative Filtering (2 points)


```{r}
df_user_191 <- read.delim("hw4.movies/user191.txt", header = FALSE, sep = "\t")
head(df_user_191)
nb_movies <- nrow(df_user_191)
```

```{r}
df_ratings <- read.csv("ml-latest-small/ratings.csv", sep=",", header=T)
df_ratings_191 <- df_ratings[which(df_ratings$userId == 191),]
df_ratings_191
```



```{r}
selected_movies <- c(150,296,380,590)
original_ratings <- integer(4)
for(i in 1:length(selected_movies)){
    rating_index <- which(paste(df_ratings_191$movieId) == selected_movies[i])
    original_ratings[i] <- df_ratings_191$rating[rating_index]
    df_ratings_191$rating[rating_index] <- NA
}
original_ratings
summary(df_ratings_191)
```

###2.3.A

```{r}
set.seed(1122)
similar_users <- c(513,317,415,375,64,556,82,225,657,266,568,50)
jaccard_sim <- c(0.43,0.40,0.32,0.30,0.275,0.272,0.25,0.24,0.226,0.221,0.21,0.20)
df_neighbors <- data.frame(userId = similar_users, similarity=jaccard_sim)
head(df_neighbors,3)
```

```{r}
random_users <- sample(df_neighbors$userId, 5)
random_users
```


```{r}
nb_movies <- nrow(df_user_191)
nb_users <- length(random_users)

list_u <- list()
list_u[[paste("userId")]] <- c(191,random_users)
for(j in 1:nb_movies){
  
  movies_rating <- integer(nb_users)
  
  id <- df_movies$movieId[which(paste(df_movies$title) == paste(df_user_191$V1[j]))]
  
  for(i in 0:nb_users+1){
  
    if(i == 1){
        user <- df_ratings[which(df_ratings$userId == 191),]
        #cat(paste("\nok"))
    }else{
        user <- df_ratings[which(df_ratings$userId == random_users[i-1]),]
        #cat(paste("\ndac"))
    }
      
    rating <- which(user$movieId == id)
    
    if(length(rating) == 0){
      movies_rating[i] <- NA
    }else{
      movies_rating[i] <- user$rating[rating]
      #cat(paste(user$rating[rating]))
    }
    
  }
  
  list_u[[paste(df_user_191$V1[j])]] <- movies_rating
}

df_u <- as.data.frame(list_u)

head(df_u)

```

```{r}
similarities <- list()
for (i in 2:6){
  #cat(paste(df_u$userId[i]," : ",which(df_neighbors$userId == df_u$userId[i]),"\n"))
  
  id <- df_u$userId[i]
  
  index <- which(df_neighbors$userId == id)
  
  result <- df_neighbors$similarity[index]
  
  similarities[[paste(id)]] <- c(id,result)
  
}

df_N <- as.data.frame(similarities)
df_n <- data.frame(t(df_N[]))
colnames(df_n) <- c("id","similarity")
df_n_sort <- df_n[order(df_n$similarity,decreasing = TRUE),]

df_n_sim <- df_n_sort[-c(4,5),]

df_n_sim
```

```{r}
tot_sim <- sum(df_n_sim$similarity)
tot_sim
```


```{r}

ratings <- integer(4)
for(i in 1:4){
  movie_id <- selected_movies[i]
  
  den <- 0
  for(j in 1:3){
    simil <-  df_n_sim$similarity[j]

    user <- df_ratings[which(df_ratings$userId == df_n_sim$id[j]),]
    rating <- user$rating[which(user$movieId == movie_id)]

    den <- den + (simil * rating)
  }
  
  new_grad <- den/tot_sim
  ratings[i] <- new_grad
  
}

ratings
```

```{r}
nb_movies <- length(selected_movies)
differences <- integer(nb_movies)
for( i in 1:nb_movies){
    cat(paste("\n",original_ratings[i],"vs :",ratings[i]))
    differences[i] <- (original_ratings[i]-ratings[i]) ** 2
}

rmse <- sqrt(sum(differences)/nb_movies)
cat(paste("\n rmse :",rmse))
```

####2.3.A-Output

```{r}
cat(paste("User ID 191, 5 random user IDs : "))
print(random_users)
cat(paste("Using user-user similarity, User ID 191 will rate the movies as follows :"))
for(i in 1:4){
  cat(paste("\n",selected_movies[i],":",round(ratings[i],1)))
}
cat(paste("\nRMSE:",trunc(rmse*10)/10))
```

###2.3.B

```{r}
df_U <- data.frame(t(df_u[-1]))
names(df_U) <- df_u[,1]

df_U$avg_mean <- apply(df_U, 1, function(x) mean(x, na.rm=T))
#df_U <- df_U[order(df_U$mean,decreasing = TRUE),]

for(i in 1:nrow(df_U)){
  #cat(paste("\n",i,": "))
  for(j in 1:6){
    #cat(paste(df_U[i,j]))
    if(!is.na(df_U[i,j])){
      df_U[i,j] <- (df_U[i,j] - df_U$avg_mean[i])
    }
    
  }
  #cat(paste("\n"))
}

original_rating_norm <- integer(4)
for(i in 1:length(selected_movies)){
  original_rating_norm[i] <- original_ratings[i] - df_U$avg_mean[i]
}

original_rating_norm

df_U <- df_U[ -c(7)]

df_U
```
```{r}
cos.sim <- function(df, index_1, index_2) 
{
    A =  as.integer(df[index_1,])
    A[is.na(A)] <- 0
    B =  as.integer(df[index_2,])
    B[is.na(B)] <- 0
    return( sum(A*B)/sqrt(sum(A^2)*sum(B^2)) )
}

cos.sim(df_U, 1, 3)

```

```{r}
nb_movies <- length(selected_movies)
ratings<- integer(nb_movies)
for( i in 1:nb_movies ) {
  
  simil <- integer(nrow(df_U))
  for(j in 1:nrow(df_U)){
    simil[j] <- cos.sim(df_U, i, j)
  }
  
  df_U$sim <- simil
  df_U <- df_U[order(df_U$sim,decreasing = TRUE),]
  df_U_top3 <- df_U[1:3,]
  
  den <- 0
  for( k in 1:3){
    den <- den + df_U_top3$sim[k] * df_U_top3$`191`[k]
    #cat(paste(j,":",den,"\n"))
  }

  rating <- den/sum(df_U_top3$sim)
  
  ratings[i] <- rating
}

ratings
```

```{r}
differences <- integer(nb_movies)
for( i in 1:nb_movies){
    cat(paste("\n",original_rating_norm[i],"vs :",ratings[i]))
    differences[i] <- (original_rating_norm[i]-ratings[i]) ** 2
}

rmse <- sqrt(sum(differences)/nb_movies)
cat(paste("\n rmse :",rmse))
```

####2.3.B-Output

```{r}
cat(paste("User ID 191, 5 random user IDs : "))
print(random_users)
cat(paste("Using item-item similarity, User ID 191 will rate the movies as follows :"))
for(i in 1:4){
  cat(paste("\n",selected_movies[i],":",round(ratings[i],1)))
}
cat(paste("\nRMSE:",round(rmse,1)))
```


